\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-06-13 paper.tex
notations

We consider a family of conditional densities $p(\mathbf{x}|\mathbf{\theta})$
defined implicitly through the simulation of a stochastic generative
process, where $\mathbf{x} \in \mathbb{R}^d$ is the observed data and
$\mathbf{\theta}$ are the parameters of interest. The simulation may involve
some complicated latent process, such that
\begin{equation}\label{eqn:p_x}
    p(\mathbf{x}|\mathbf{\theta}) = \int p(\mathbf{x}|\mathbf{z},\mathbf{\theta}) p(\mathbf{z}) d\mathbf{z}
\end{equation}
where $\mathbf{z} \in \mathbb{R}^m$ is a latent variable providing an external source
of randomness.

We assume that we already have an accurate simulation of the stochastic generative process
$p(\mathbf{x}|\mathbf{z},\mathbf{\theta})$, as specified through a deterministic function $g(\cdot;
\theta) : \mathbb{R}^m \to \mathbb{R}^d$. That is,
\begin{equation}\label{eqn:p_x_sim}
    p(\mathbf{x}|\mathbf{\theta}) = \frac{\partial}{\partial x_1} \dots \frac{\partial}{\partial x_d} \int_{\{\mathbf{z}:g(\mathbf{z};\mathbf{\theta}) \leq \mathbf{x}\}} p(\mathbf{z}) d\mathbf{z}.
\end{equation}
The simulator $g$ is assumed to be a non-invertible function, that can
only be used to generate data in forward mode.
For this reason, evaluating the integral in Eqn.~\ref{eqn:p_x_sim} is intractable.
Importantly, and as increasingly found in science, we consider the additional constraint that $g$ is a non-differentiable
model, e.g. as when specified as a computer program.

Given some observed data $\{ \mathbf{x}_i | i=1, \dots, N \}$, our goal is the inference
of the parameters of interest $\mathbf{\theta}^*$ that maximize the (marginal) log-likelihood of the observations:
\begin{equation}
    \mathbf{\theta}^* = \arg \max_\mathbf{\theta} \sum_i \log p(\mathbf{x}_i | \mathbf{\theta}).
\end{equation}
\glnote{Redefine the goal as the minimization of a divergence between the true and the induced distribution?}

@@

Generative adversarial networks were first proposed by
\cite{goodfellow2014generative} as a way to build an implicit generative model
capable of producing samples from random noise $\mathbf{z}$. More specifically,
a generative model $g(\cdot; \mathbf{\theta})$ is pit against an adversarial
network $d$ whose antigonistic objective is to recognize real data $\mathbf{x}$
from generated data $g(\mathbf{z}; \mathbf{\theta})$. Both models $g$ and $d$
are trained simultaneously, in such a way that $g$ learns to maximally confuse
its adversary $d$ (which happens when $g$ produces samples comparable to the
observed data), while $d$ continuously adapts to changes in $g$. At the
equilibrium and assuming enough capacity in the networks, it can be shown that
$g$ induces a distribution  that is indistinguishable from the distribution of
the observed data $\mathbf{x}$.

>>>>

We consider a family of parameterized densities $p_\mathbf{\theta}(\mathbf{x})$
defined implicitly through the simulation of a stochastic generative
process, where $\mathbf{x} \in \mathbb{R}^d$ is the data and
$\mathbf{\theta}$ are the parameters of interest. The simulation may involve
some complicated latent process, such that
\begin{equation}\label{eqn:p_x}
    p_\mathbf{\theta}(\mathbf{x}) = \int p_\mathbf{\theta}(\mathbf{x}|\mathbf{z}) p(\mathbf{z}) d\mathbf{z}
\end{equation}
where $\mathbf{z} \in \mathbb{R}^m$ is a latent variable providing an external source
of randomness.

We assume that we already have an accurate simulation of the stochastic generative process
that defines $p_\mathbf{\theta}(\mathbf{x}|\mathbf{z})$, as specified through a deterministic function $g(\cdot;
\theta) : \mathbb{R}^m \to \mathbb{R}^d$. That is,
\begin{equation}\label{eqn:p_x_sim}
    p_\mathbf{\theta}(\mathbf{x}) = \frac{\partial}{\partial x_1} \dots \frac{\partial}{\partial x_d} \int_{\{\mathbf{z}:g(\mathbf{z};\mathbf{\theta}) \leq \mathbf{x}\}} p(\mathbf{z}) d\mathbf{z}.
\end{equation}
The simulator $g$ is assumed to be a non-invertible function, that can
only be used to generate data in forward mode.
For this reason, evaluating the integral in Eqn.~\ref{eqn:p_x_sim} is intractable.
Importantly, and as increasingly found in science, we consider the additional constraint that $g$ is a non-differentiable
model, e.g. as when specified as a computer program.

Given some observed data $\{ \mathbf{x}_i | i=1, \dots, N \}$ drawn from the (unknown) true distribution $p_r$, our goal is the inference
of the parameters of interest $\mathbf{\theta}^*$ that minimize the divergence
between $p_r$ and the modeled
data distribution $p_\mathbf{\theta}$ induced by $g(\cdot; \mathbf{\theta})$. That is,
\begin{equation}
    \mathbf{\theta}^* = \arg \min_\mathbf{\theta} \rho(p_r, p_\mathbf{\theta}),
\end{equation}
where $\rho$ is some divergence.

@@

Generative adversarial networks (GAN) were first proposed by
\cite{goodfellow2014generative} as a way to build an implicit generative model
capable of producing samples from random noise $\mathbf{z}$. More specifically,
a generative model $g(\cdot; \mathbf{\theta})$ is pit against an adversarial
network $d$ whose antigonistic objective is to recognize real data $\mathbf{x}$
from generated data $g(\mathbf{z}; \mathbf{\theta})$. Both models $g$ and $d$
are trained simultaneously, in such a way that $g$ learns to maximally confuse
its adversary $d$ (which happens when $g$ produces samples comparable to the
observed data), while $d$ continuously adapts to changes in $g$. At the
equilibrium and assuming enough capacity in the networks, it can be shown that
$g$ induces a distribution  that is indistinguishable from the distribution of
the observed data $\mathbf{x}$.

More recently, ...
